### 2021-02-19 ch05_paper_ecs/paper_ecs.tex
Added sections 6.2.1 and 6.2.2

In this \namecref{sec:05:comparison_of_emergent_constraints} we describe and
discuss the eleven emergent constraints on \ac{ECS} summarized in
\cref{tab:05:overview_emergent_constraints} using \acs{CMIP}5 and \acs{CMIP}6
data
(\cref{subsec:05:bri,subsec:05:cox,subsec:05:lip,subsec:05:shd,subsec:05:shs,%
  subsec:05:shl,subsec:05:su,subsec:05:tih,subsec:05:tii,subsec:05:vol,%
  subsec:05:zha}) and provide a best estimate for \ac{ECS} and statistical
significance of the eleven emergent constraints in
\cref{subsec:05:emergent_constraints_summary}. While most of these emergent
constraints have been derived using data from the \acs{CMIP}5 and/or
\acs{CMIP}3 ensembles, to our knowledge none of them has been evaluated on
the \acs{CMIP}6 ensemble so far. The results for the individual emergent
constraints described in the following are shown in
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}.
The left columns in these figures show the emergent relationships, including
the uncertainty of the linear regressions (blue and orange shaded areas; see
\cref{eq:02:spe}) and the uncertainty in the observations (gray shaded area;
see \cref{eq:02:pdf_obs}). The right columns show the \acp{PDF} of \ac{ECS} in
the original model ensemble (histogram) and the constrained distribution given
by the emergent constraints (blue and orange line; see
\cref{eq:02:pdf_y_given_x0}). \Cref{tab:05:overview_results} shows the
corresponding $66 \unit{\%}$ confidence intervals (\ie{} the $\range{17}{83}
\unit{\%}$ intervals) of \ac{ECS} derived from the \acp{PDF} given by
\cref{eq:02:pdf_y_given_x0} and the $p$-values used to assess the significance
of the emergent relationships.

@@

\textcite{Tian2015} finds a link between mid-tropospheric humidity over the
tropical Pacific and simulated moisture, precipitation, clouds, large-scale
circulation, and thus \ac{ECS} in \acs{CMIP}3 and \acs{CMIP}5 models. The study
explains this link with the similarity of mid-tropospheric humidity and
precipitation patterns as both are related to the \ac{ITCZ}. The proposed
tropical mid-tropospheric humidity asymmetry index to constrain \ac{ECS} is
defined as relative bias (in $\si{\%}$) in simulated annual mean $500
\unit{hPa}$ specific humidity averaged over the Southern Hemisphere tropical
Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$, $\range{80
  \unit{\degree W}}{120 \unit{\degree E}}$) minus the bias averaged over the
Northern Hemisphere tropical Pacific ($\range{0 \unit{\degree N}}{20
  \unit{\degree N}}$, $\range{80 \unit{\degree W}}{120 \unit{\degree E}}$) when
compared with observations. Similar to the SU constraint, the index proposed by
\textcite{Tian2015} seems to be related to the emergent constraint by
\textcite{Fasullo2012}, who find correlations between relative humidity of the
middle and upper troposphere and \ac{ECS}. Here, we use humidity observations
from AIRS \autocite{Aumann2003} over the time period \range{2003}{2005} as the
reference dataset. We assess a $66 \unit{\%}$ \ac{ECS} range of
$\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5 ($R^2 = 0.24$) and
$\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 = 0.06$).
\textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
significance of the emergent relationship drops massively from $p = 0.0089$ in
\acs{CMIP}5 to $p = 0.1348$ in \acs{CMIP}6.

@@

In addition to the humidity index, \textcite{Tian2015} proposes an emergent
constraint on \ac{ECS} based on the southern \ac{ITCZ} index
\autocite{Bellucci2010, Hirota2011}. This index is defined as the
climatological annual mean precipitation bias averaged over the southeastern
Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$, $\range{150
  \unit{\degree W}}{100 \unit{\degree W}}$). The southern \ac{ITCZ} index is
calculated in $\si{\mmday{}}$ and dominated by the so-called double-\ac{ITCZ},
a common problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds
a link between double-\ac{ITCZ} bias and simulated moisture, precipitation,
clouds, and large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He
argues that this could explain the link between the double-\acs{ITCZ} bias and
\ac{ECS}. As reference data, we use observed precipitation data for the years
\range{1986}{2005} from GPCP \autocite{Adler2003}. We calculate an \ac{ECS} $66
\unit{\%}$ confidence range of $\pmrangeunit{3.87}{0.67}{K}$ for \acs{CMIP}5
($R^2 = 0.33$) and $\pmrangeunit{3.84}{1.09}{K}$ for \acs{CMIP}6 ($R^2 <
0.01$). \textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
emergent relationship shows a much lower statistical significance in
\acs{CMIP}6 ($p = 0.8236$) than in \acs{CMIP}5 ($p = 0.0013$).

@@

The study by \textcite{Volodin2008} aims at constraining \ac{ECS} based on the
geographical distribution of clouds in climate models. Since this early
emergent constraint has originally been trained on \acs{CMIP}3 models, both
\acs{CMIP}5 and \acs{CMIP}6 are out-of-sample tests for it.
\textcite{Volodin2008} shows that high \ac{ECS} models tend to simulate a
higher total cloud cover over the southern midlatitudes and a lower total cloud
cover over the tropics (relative to the \ac{MMM}). This can be used to
establish an emergent relationship between the \ac{ECS} and the difference in
tropical total cloud cover ($\range{28 \unit{\degree S}}{28 \unit{\degree N}}$)
and the southern midlatitudinal total cloud cover ($\range{56 \unit{\degree
    S}}{36 \unit{\degree S}}$). Analogous to the original study, we use the
ISCCP-D2 data \autocite{Rossow1991} as observational reference. For the VOL
constraint, we calculate a constrained $66 \unit{\%}$ range of \ac{ECS} of
$\pmrangeunit{3.74}{0.64}{K}$ for \acs{CMIP}5 ($R^2 = 0.38$) and
$\pmrangeunit{4.21}{1.04}{K}$ for \acs{CMIP}6 ($R^2 = 0.18$), whereas the
original publication gives a range of $\pmrangeunit{3.6}{0.4}{K}$ (standard
deviation) for a climate model ensemble of \acs{CMIP}3 models. The emergent
constraint by \textcite{Volodin2008} shows a lower significance in the
\acs{CMIP}6 ensemble ($p = 0.0056$) than in the \acs{CMIP}5 ensemble ($p =
0.0004$).

@@

\textcite{Zhai2015} focus on the variations of \acp{MBLC}, which largely
contribute to the shortwave cloud feedback and thus to the uncertainty in
modeled \ac{ECS}. Their central quantity is the response of the \ac{MBLC}
fraction to changes in the \ac{SST} in subtropical oceanic subsidence regions
for both hemispheres ($\range{20 \unit{\degree}}{40 \unit{\degree}}$). On short
(seasonal) and long (centennial under a forcing) time scales, this quantity is
well correlated with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5
models. Together with observations of cloud fraction from Cloudsat/CALIPSO
\autocite{Mace2009}, \ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and
vertical velocity from ERA-Interim \autocite{Dee2011}, the seasonal response of
\ac{MBLC} fraction to changes in \ac{SST} forms an emergent constraint on
\ac{ECS}. We assess a $66 \unit{\%}$ \ac{ECS} range of
$\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5 ($R^2 = 0.05$) and
$\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 = 0.62$). In their original
publication, \textcite{Zhai2015} find an \ac{ECS} range of
$\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination of
\acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance, the
results of the ZHA constraints are somewhat surprising: although \acs{CMIP}5
data (in combination with \acs{CMIP}3 data) has successfully been used in the
original publication, our approach finds that the statistical significance of
the emergent relationship is much higher in the unseen \acs{CMIP}6 ensemble ($p
< 0.0001$) than in the previously available \acs{CMIP}5 ensemble ($p =
0.2567$). The ZHA constraint is the only emergent constraint analyzed here that
shows this extreme behavior (only one other constraint, SHS, shows a slightly
higher significance in \acs{CMIP}6; all other constraints show lower
significances in \acs{CMIP}6). The reason for the erratic skill in \acs{CMIP}5
is the set of climate models used. For our analysis, we use eleven additional
\acs{CMIP}5 models that have not been used in the original publication
(ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m, CCSM4, GFDL-ESM2G, GFDL-ESM2M,
IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and MPI-ESM-P). Due to a lack of
publicly available data, the model CESM1-CAM5 that is used in the original
publication is not included in our analysis. The effect of choosing different
subsets of \acs{CMIP}5 models on the emergent relationship is illustrated in
\cref{fig:05:zha}. Using the original \acs{CMIP}5 models from the original
publication gives a considerably higher correlation ($R^2 = 0.38$) than using
all available \acs{CMIP}5 models ($R^2 = 0.05$). This result shows a strong
dependency of this emergent constraint on the subset of climate models used.
Nonetheless, the performance on \acs{CMIP}6 models is, surprisingly, the best
of all the constraints and much better than on either subset of \acs{CMIP}5
models.

@@

\begin{table}[!b]
  \centering
  \begin{tabular}{p{0.06\columnwidth} p{0.34\columnwidth} y{0.12} y{0.12}
      y{0.1} y{0.1}}
    \toprule
    Label & \ecreshead{\acs{ECS}}{(original publication)} &
    \ecreshead{\acs{ECS} [$\si{K}$]}{(\acs{CMIP}5)} & \ecreshead{\acs{ECS}
      [$\si{K}$]}{(\acs{CMIP}6)} & \ecreshead{$p$}{(\acs{CMIP}5)} &
    \ecreshead{$p$}{(\acs{CMIP}6)} \\
    \midrule
    BRI & most likely $4.0 \unit{K}$, $< 2.30 \unit{K}$ very unlikely ($90
    \unit{\%}$ confidence) & 3.72
    \pm 0.59 & 4.32 \pm 1.07 & 0.0005 & 0.0355 \\
    COX & $\pmrangeunit{2.8}{0.6}{K}$ & 3.03 \pm 0.73 & 3.71 \pm 1.09 & 0.0032
    & 0.5415 \\
    LIP & no best estimate given & 2.97 \pm 0.75 & 3.75 \pm 1.11 & 0.0228 &
    0.6791 \\
    SHD & none --- see SHL & 3.65 \pm 0.64 & 3.77 \pm 1.06 & 0.0037 & 0.2805 \\
    SHL & most likely $4 \unit{K}$, lower limit $3 \unit{K}$ & 3.42 \pm 0.65 &
    3.67 \pm 1.06 & 0.0002 & 0.0138 \\
    SHS & none --- see SHL & 3.07 \pm 0.73 & 3.48 \pm 1.07 & 0.0647 & 0.0396 \\
    SU & most likely $4 \unit{K}$, lower limit $3 \unit{K}$ & 3.30 \pm 0.88 &
    3.77 \pm 1.35 & 0.1676 & 0.1935 \\
    TIH & most likely $4.0 \unit{K}$ & 3.88 \pm 0.75 & 4.15 \pm 1.10 & 0.0089
    & 0.1348 \\
    TII & most likely $4.0 \unit{K}$ & 3.87 \pm 0.67 & 3.84 \pm 1.09 & 0.0013
    & 0.8236 \\
    VOL & $\pmrangeunit{3.6}{0.4}{K}$ (\stddev{}) & 3.74 \pm 0.64 & 4.21 \pm
    1.04 & 0.0004 & 0.0056 \\
    ZHA & $\pmrangeunit{3.90}{0.45}{K}$ (\stddev{}) & 3.35 \pm 0.74 & 3.79 \pm
    0.67 & 0.2567 & < 0.0001 \\
    \bottomrule
  \end{tabular}
  \caption{Overview of the constrained \acf{ECS} ranges and $p$-values for all
    eleven analyzed emergent constraints. If not further specified, the
    uncertainty ranges correspond to the $66 \unit{\%}$ confidence intervals
    ($\range{17}{83} \unit{\%}$). For \acs{CMIP}5 and \acs{CMIP}6, these are
    evaluated from the \acl{PDF} given by \cref{eq:02:pdf_y_given_x0} (see also
    the right columns of \cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,%
      fig:05:su_tih_tii,fig:05:vol_zha}). Note that even though \acs{CMIP}5
    models have been used for some constraints in the original publications,
    the constrained ranges in the second and third column might differ due to
    the use of a different subset of climate models (in this study, we use
    output from all \acs{CMIP} models that is publicly available; see
    \cref{tab:app:a:cmip5_emergent_constraints_part1,%
      tab:app:a:cmip5_emergent_constraints_part2,%
      tab:app:a:cmip6_emergent_constraints_part1,%
      tab:app:a:cmip6_emergent_constraints_part2} for details). The $p$-values
    describing the significance of the emergent relationships are defined as
    the probability to obtain an absolute correlation coefficient $\abs*{r}$
    or higher under the null hypothesis that the true underlying correlation
    coefficient between the predictor and \acs{ECS} is zero. Smaller
    $p$-values point to higher significance and vice versa (see
    \cref{subsec:02:emergent_constraints} for details).
    \AdaptedFrom{Schlund2020a}.}
  \label{tab:05:overview_results}
\end{table}

In most cases, the emergent relationships (left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha})
show the same sign of the slope (as expected from the theory) for \acs{CMIP}5
and \acs{CMIP}6, with the SU constraint being the only exception. However, the
coefficient of determination ($R^2$) is lower for \acs{CMIP}6 compared to
\acs{CMIP}5 for all one constraint: ZHA. The probability distributions of the
constrained \ac{ECS} that we obtain (right columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha})
give similar results: except for the ZHA constraint, the constraint on the
\acs{CMIP}6 ensemble is weaker, \ie{} the constrained \acp{PDF} derived from
the \acs{CMIP}6 ensemble are broader than their respective \acs{CMIP}5
counterparts. As shown in \cref{tab:05:overview_results} and
\cref{fig:05:overview_results}, for \acs{CMIP}5, the range of the best (maximum
likelihood) estimates for \acs{ECS} is $2.97 \unit{K}$ to $3.88 \unit{K}$,
while the corresponding \acs{CMIP}6 best estimates are higher for almost every
tested emergent constraint (TII being the only exception), resulting in a range
of best estimates of $3.48 \unit{K}$ to $4.32 \unit{K}$. Using the arithmetic
mean of all analyzed emergent constraints, this results in a mean increase of
the \ac{ECS} best estimate of $12 \unit{\%}$ in \acs{CMIP}6 compared to
\acs{CMIP}5. Similarly, the size of the $66 \unit{\%}$ \ac{ECS} ranges
($\range{17}{83} \unit{\%}$ confidence) shows values of $1.16 \unit{K}$ to
$1.75 \unit{K}$ in \acs{CMIP}5 and $1.32 \unit{K}$ to $2.70 \unit{K}$ in
\acs{CMIP}6, resulting in an increase of $51 \unit{\%}$ averaged over all
emergent constraints.

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch05_paper_ecs/figs/constrained_ranges_summary.pdf}
  \caption{Overview of the constrained \acf{ECS} ranges given by the eleven
    analyzed emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6
    (orange) ensemble. The bars correspond to the $66 \unit{\%}$ confidence
    intervals ($\range{17}{83} \unit{\%}$) calculated from the constrained
    \acl{PDF} of \acs{ECS} following \cref{eq:02:pdf_y_given_x0}. The vertical
    black lines illustrate the corresponding best estimates of the
    distributions. For almost all emergent constraints, the constrained best
    estimate and range of \ac{ECS} is higher in the \acs{CMIP}6 ensemble.}
  \label{fig:05:overview_results}
\end{figure}

>>>>

In this \namecref{sec:05:comparison_of_emergent_constraints} we describe and
discuss the eleven emergent constraints on \ac{ECS} summarized in
\cref{tab:05:overview_emergent_constraints} using \acs{CMIP}5 and \acs{CMIP}6
data
(\cref{subsec:05:bri,subsec:05:cox,subsec:05:lip,subsec:05:shd,subsec:05:shs,%
  subsec:05:shl,subsec:05:su,subsec:05:tih,subsec:05:tii,subsec:05:vol,%
  subsec:05:zha}) and provide a best estimate for \ac{ECS} and statistical
significance of the eleven emergent constraints in
\cref{subsec:05:emergent_constraints_summary}. While most of these emergent
constraints have been derived using data from the \acs{CMIP}5 and/or
\acs{CMIP}3 ensembles, to our knowledge none of them has been evaluated on
the \acs{CMIP}6 ensemble so far. The results for the individual emergent
constraints described in the following are shown in
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}.
The left columns in these figures show the emergent relationships, including
the uncertainty of the linear regressions (blue and orange shaded areas; see
\cref{eq:02:spe}) and the uncertainty in the observations (gray shaded area;
see \cref{eq:02:pdf_obs}). The right columns show the \acp{PDF} of \ac{ECS} in
the original model ensemble (histogram) and the constrained distribution given
by the emergent constraints (blue and orange line; see
\cref{eq:02:pdf_y_given_x0}). \Cref{tab:05:overview_results} shows the
corresponding $66 \unit{\%}$ confidence intervals (\ie{} the
$\rangeunit{17}{83}{\%}$ intervals) of \ac{ECS} derived from the \acp{PDF}
given by \cref{eq:02:pdf_y_given_x0} and the $p$-values used to assess the
significance of the emergent relationships.

@@

\textcite{Tian2015} finds a link between mid-tropospheric humidity over the
tropical Pacific and simulated moisture, precipitation, clouds, large-scale
circulation, and thus \ac{ECS} in \acs{CMIP}3 and \acs{CMIP}5 models. The study
explains this link with the similarity of mid-tropospheric humidity and
precipitation patterns as both are related to the \ac{ITCZ}. The proposed
tropical mid-tropospheric humidity asymmetry index to constrain \ac{ECS} is
defined as relative bias (in $\si{\%}$) in simulated annual mean $500
\unit{hPa}$ specific humidity averaged over the Southern Hemisphere tropical
Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$, $\range{80
  \unit{\degree W}}{120 \unit{\degree E}}$) minus the bias averaged over the
Northern Hemisphere tropical Pacific ($\rangeunit{0}{20}{\degree N}$,
$\range{80 \unit{\degree W}}{120 \unit{\degree E}}$) when compared with
observations. Similar to the SU constraint, the index proposed by
\textcite{Tian2015} seems to be related to the emergent constraint by
\textcite{Fasullo2012}, who find correlations between relative humidity of the
middle and upper troposphere and \ac{ECS}. Here, we use humidity observations
from AIRS \autocite{Aumann2003} over the time period \range{2003}{2005} as the
reference dataset. We assess a $66 \unit{\%}$ \ac{ECS} range of
$\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5 ($R^2 = 0.24$) and
$\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 = 0.06$).
\textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
significance of the emergent relationship drops massively from $p = 0.0089$ in
\acs{CMIP}5 to $p = 0.1348$ in \acs{CMIP}6.

@@

In addition to the humidity index, \textcite{Tian2015} proposes an emergent
constraint on \ac{ECS} based on the southern \ac{ITCZ} index
\autocite{Bellucci2010, Hirota2011}. This index is defined as the
climatological annual mean precipitation bias averaged over the southeastern
Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$,
$\rangeunit{150}{100}{\degree W}$). The southern \ac{ITCZ} index is calculated
in $\si{\mmday{}}$ and dominated by the so-called double-\ac{ITCZ}, a common
problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds a link
between double-\ac{ITCZ} bias and simulated moisture, precipitation, clouds,
and large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He argues
that this could explain the link between the double-\acs{ITCZ} bias and
\ac{ECS}. As reference data, we use observed precipitation data for the years
\range{1986}{2005} from GPCP \autocite{Adler2003}. We calculate an \ac{ECS} $66
\unit{\%}$ confidence range of $\pmrangeunit{3.87}{0.67}{K}$ for \acs{CMIP}5
($R^2 = 0.33$) and $\pmrangeunit{3.84}{1.09}{K}$ for \acs{CMIP}6 ($R^2 <
0.01$). \textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
emergent relationship shows a much lower statistical significance in
\acs{CMIP}6 ($p = 0.8236$) than in \acs{CMIP}5 ($p = 0.0013$).

@@

The study by \textcite{Volodin2008} aims at constraining \ac{ECS} based on the
geographical distribution of clouds in climate models. Since this early
emergent constraint has originally been trained on \acs{CMIP}3 models, both
\acs{CMIP}5 and \acs{CMIP}6 are out-of-sample tests for it.
\textcite{Volodin2008} shows that high \ac{ECS} models tend to simulate a
higher total cloud cover over the southern midlatitudes and a lower total cloud
cover over the tropics (relative to the \ac{MMM}). This can be used to
establish an emergent relationship between the \ac{ECS} and the difference in
tropical total cloud cover ($\range{28 \unit{\degree S}}{28 \unit{\degree N}}$)
and the southern midlatitudinal total cloud cover ($\rangeunit{56}{36}{\degree
  S}$). Analogous to the original study, we use the ISCCP-D2 data
\autocite{Rossow1991} as observational reference. For the VOL constraint, we
calculate a constrained $66 \unit{\%}$ range of \ac{ECS} of
$\pmrangeunit{3.74}{0.64}{K}$ for \acs{CMIP}5 ($R^2 = 0.38$) and
$\pmrangeunit{4.21}{1.04}{K}$ for \acs{CMIP}6 ($R^2 = 0.18$), whereas the
original publication gives a range of $\pmrangeunit{3.6}{0.4}{K}$ (standard
deviation) for a climate model ensemble of \acs{CMIP}3 models. The emergent
constraint by \textcite{Volodin2008} shows a lower significance in the
\acs{CMIP}6 ensemble ($p = 0.0056$) than in the \acs{CMIP}5 ensemble ($p =
0.0004$).

@@

\textcite{Zhai2015} focus on the variations of \acp{MBLC}, which largely
contribute to the shortwave cloud feedback and thus to the uncertainty in
modeled \ac{ECS}. Their central quantity is the response of the \ac{MBLC}
fraction to changes in the \ac{SST} in subtropical oceanic subsidence regions
for both hemispheres ($\rangeunit{20}{40}{\degree}$). On short (seasonal) and
long (centennial under a forcing) time scales, this quantity is well correlated
with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5 models. Together
with observations of cloud fraction from Cloudsat/CALIPSO \autocite{Mace2009},
\ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and vertical velocity from
ERA-Interim \autocite{Dee2011}, the seasonal response of \ac{MBLC} fraction to
changes in \ac{SST} forms an emergent constraint on \ac{ECS}. We assess a $66
\unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5
($R^2 = 0.05$) and $\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 =
0.62$). In their original publication, \textcite{Zhai2015} find an \ac{ECS}
range of $\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination
of \acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance,
the results of the ZHA constraints are somewhat surprising: although
\acs{CMIP}5 data (in combination with \acs{CMIP}3 data) has successfully been
used in the original publication, our approach finds that the statistical
significance of the emergent relationship is much higher in the unseen
\acs{CMIP}6 ensemble ($p < 0.0001$) than in the previously available
\acs{CMIP}5 ensemble ($p = 0.2567$). The ZHA constraint is the only emergent
constraint analyzed here that shows this extreme behavior (only one other
constraint, SHS, shows a slightly higher significance in \acs{CMIP}6; all other
constraints show lower significances in \acs{CMIP}6). The reason for the
erratic skill in \acs{CMIP}5 is the set of climate models used. For our
analysis, we use eleven additional \acs{CMIP}5 models that have not been used
in the original publication (ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m,
CCSM4, GFDL-ESM2G, GFDL-ESM2M, IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and
MPI-ESM-P). Due to a lack of publicly available data, the model CESM1-CAM5 that
is used in the original publication is not included in our analysis. The effect
of choosing different subsets of \acs{CMIP}5 models on the emergent
relationship is illustrated in \cref{fig:05:zha}. Using the original
\acs{CMIP}5 models from the original publication gives a considerably higher
correlation ($R^2 = 0.38$) than using all available \acs{CMIP}5 models ($R^2 =
0.05$). This result shows a strong dependency of this emergent constraint on
the subset of climate models used. Nonetheless, the performance on \acs{CMIP}6
models is, surprisingly, the best of all the constraints and much better than
on either subset of \acs{CMIP}5 models.

@@

\begin{table}[!b]
  \centering
  \begin{tabular}{p{0.06\columnwidth} p{0.34\columnwidth} y{0.12} y{0.12}
      y{0.1} y{0.1}}
    \toprule
    Label & \ecreshead{\acs{ECS}}{(original publication)} &
    \ecreshead{\acs{ECS} [$\si{K}$]}{(\acs{CMIP}5)} & \ecreshead{\acs{ECS}
      [$\si{K}$]}{(\acs{CMIP}6)} & \ecreshead{$p$}{(\acs{CMIP}5)} &
    \ecreshead{$p$}{(\acs{CMIP}6)} \\
    \midrule
    BRI & most likely $4.0 \unit{K}$, $< 2.30 \unit{K}$ very unlikely ($90
    \unit{\%}$ confidence) & 3.72
    \pm 0.59 & 4.32 \pm 1.07 & 0.0005 & 0.0355 \\
    COX & $\pmrangeunit{2.8}{0.6}{K}$ & 3.03 \pm 0.73 & 3.71 \pm 1.09 & 0.0032
    & 0.5415 \\
    LIP & no best estimate given & 2.97 \pm 0.75 & 3.75 \pm 1.11 & 0.0228 &
    0.6791 \\
    SHD & none --- see SHL & 3.65 \pm 0.64 & 3.77 \pm 1.06 & 0.0037 & 0.2805 \\
    SHL & most likely $4 \unit{K}$, lower limit $3 \unit{K}$ & 3.42 \pm 0.65 &
    3.67 \pm 1.06 & 0.0002 & 0.0138 \\
    SHS & none --- see SHL & 3.07 \pm 0.73 & 3.48 \pm 1.07 & 0.0647 & 0.0396 \\
    SU & most likely $4 \unit{K}$, lower limit $3 \unit{K}$ & 3.30 \pm 0.88 &
    3.77 \pm 1.35 & 0.1676 & 0.1935 \\
    TIH & most likely $4.0 \unit{K}$ & 3.88 \pm 0.75 & 4.15 \pm 1.10 & 0.0089
    & 0.1348 \\
    TII & most likely $4.0 \unit{K}$ & 3.87 \pm 0.67 & 3.84 \pm 1.09 & 0.0013
    & 0.8236 \\
    VOL & $\pmrangeunit{3.6}{0.4}{K}$ (\stddev{}) & 3.74 \pm 0.64 & 4.21 \pm
    1.04 & 0.0004 & 0.0056 \\
    ZHA & $\pmrangeunit{3.90}{0.45}{K}$ (\stddev{}) & 3.35 \pm 0.74 & 3.79 \pm
    0.67 & 0.2567 & < 0.0001 \\
    \bottomrule
  \end{tabular}
  \caption{Overview of the constrained \acf{ECS} ranges and $p$-values for all
    eleven analyzed emergent constraints. If not further specified, the
    uncertainty ranges correspond to the $66 \unit{\%}$ confidence intervals
    ($\rangeunit{17}{83}{\%}$). For \acs{CMIP}5 and \acs{CMIP}6, these are
    evaluated from the \acl{PDF} given by \cref{eq:02:pdf_y_given_x0} (see also
    the right columns of \cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,%
      fig:05:su_tih_tii,fig:05:vol_zha}). Note that even though \acs{CMIP}5
    models have been used for some constraints in the original publications,
    the constrained ranges in the second and third column might differ due to
    the use of a different subset of climate models (in this study, we use
    output from all \acs{CMIP} models that is publicly available; see
    \cref{tab:app:a:cmip5_emergent_constraints_part1,%
      tab:app:a:cmip5_emergent_constraints_part2,%
      tab:app:a:cmip6_emergent_constraints_part1,%
      tab:app:a:cmip6_emergent_constraints_part2} for details). The $p$-values
    describing the significance of the emergent relationships are defined as
    the probability to obtain an absolute correlation coefficient $\abs*{r}$
    or higher under the null hypothesis that the true underlying correlation
    coefficient between the predictor and \acs{ECS} is zero. Smaller
    $p$-values point to higher significance and vice versa (see
    \cref{subsec:02:emergent_constraints} for details).
    \AdaptedFrom{Schlund2020a}.}
  \label{tab:05:overview_results}
\end{table}

In most cases, the emergent relationships (left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha})
show the same sign of the slope (as expected from the theory) for \acs{CMIP}5
and \acs{CMIP}6, with the SU constraint being the only exception. However, the
coefficient of determination ($R^2$) is lower for \acs{CMIP}6 compared to
\acs{CMIP}5 for all one constraint: ZHA. The probability distributions of the
constrained \ac{ECS} that we obtain (right columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha})
give similar results: except for the ZHA constraint, the constraint on the
\acs{CMIP}6 ensemble is weaker, \ie{} the constrained \acp{PDF} derived from
the \acs{CMIP}6 ensemble are broader than their respective \acs{CMIP}5
counterparts. As shown in \cref{tab:05:overview_results} and
\cref{fig:05:overview_results}, for \acs{CMIP}5, the range of the best (maximum
likelihood) estimates for \acs{ECS} is $2.97 \unit{K}$ to $3.88 \unit{K}$,
while the corresponding \acs{CMIP}6 best estimates are higher for almost every
tested emergent constraint (TII being the only exception), resulting in a range
of best estimates of $3.48 \unit{K}$ to $4.32 \unit{K}$. Using the arithmetic
mean of all analyzed emergent constraints, this results in a mean increase of
the \ac{ECS} best estimate of $12 \unit{\%}$ in \acs{CMIP}6 compared to
\acs{CMIP}5. Similarly, the size of the $66 \unit{\%}$ \ac{ECS} ranges
($\rangeunit{17}{83}{\%}$ confidence) shows values of $1.16 \unit{K}$ to $1.75
\unit{K}$ in \acs{CMIP}5 and $1.32 \unit{K}$ to $2.70 \unit{K}$ in \acs{CMIP}6,
resulting in an increase of $51 \unit{\%}$ averaged over all emergent
constraints.

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch05_paper_ecs/figs/constrained_ranges_summary.pdf}
  \caption{Overview of the constrained \acf{ECS} ranges given by the eleven
    analyzed emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6
    (orange) ensemble. The bars correspond to the $66 \unit{\%}$ confidence
    intervals ($\rangeunit{17}{83}{\%}$) calculated from the constrained
    \acl{PDF} of \acs{ECS} following \cref{eq:02:pdf_y_given_x0}. The vertical
    black lines illustrate the corresponding best estimates of the
    distributions. For almost all emergent constraints, the constrained best
    estimate and range of \ac{ECS} is higher in the \acs{CMIP}6 ensemble.}
  \label{fig:05:overview_results}
\end{figure}
